### 2023-07-25 literature_review.tex

\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}

\usepackage{amsmath,amsfonts,amssymb,amsthm,
mathtools,mathrsfs}
\usepackage{physics}
\usepackage{xcolor}
\usepackage{hyperref}
% Citation style
\usepackage[style=apa, backend=biber, natbib]{biblatex}
\addbibresource{references.bib}

\title{Literature review}
\author{Mingxuan He}

\begin{document}

\maketitle

\section*{Micro foundations}

\begin{itemize}
    \item \cite{nisan2007algorithmic}, etc. Algorithmic game theory
    \item \cite{gans2022mechanism}, etc. Mechanism design
    \item \cite{budish2018economic, gans2019more} Costs of securing blockchain
\end{itemize}

\section*{Pricing models of cryptocurrencies}
\begin{itemize}
    \item \cite{yermack2015bitcoin} Bitcoin is not a currency
    \item \cite{cong2021tokenomics, cong2022token} Asset pricing model featuring token issuance as means of platform financing and user growth
    \item \cite{athey2016bitcoin, catalini2020some} Partial equilibrium
    \item \cite{schilling2019some, schilling2019currency} Crypto vs. Fiat as mediums of exchange with transaction fees
    \item \cite{catalini2020markets} Demand and supply of PoS tokens under attack
    \item \cite{biais2020equilibrium} An OLG model of prices and transactional benefits
    \item \cite{bolt2020value} Modeled crypto exchange rates with quantity theory of money
\end{itemize}

\section*{Monetary policies for stablecoins, platform tokens, and CBDCs}

\end{document}