### 2023-09-23 Figures/flowchart_baseline.tex
move flowcharts

\documentclass[../Paper/main.tex]{subfiles}

\begin{document}

\begin{equation}
    \label{fig:tikz-b}
    \begin{tikzcd}[ampersand replacement=\&, column sep = 4.5em, row sep = 4.5em]
    Protocol 
    \ar[d, "\substack{\text{newly minted}\\\text{tokens}}" ']
    \&
    \ar[d, shift right, "\substack{\text{endow-}\\ \text{ment}}" 'near start]
    \ar[d, shift left, "\substack{\text{purchase}\\ \text{assets}}" near start]
    \&
    \ar[d, "\substack{\text{transact.}\\ \text{benefits}}" ]
    \&
    \\
    Validators
    \& 
    \substack{Young\\Users}
    \ar[l,"\substack{\text{transact.}\\ \text{fees}}" ']
    \ar[r, dashed, "t+1", 
    "\substack{\text{hold}\\\text{assets}}" ']
    \& 
    \substack{Old\\Users}
    \ar[r, "\substack{\text{hacked}\\ \text{tokens}}"]
    \&
    Hackers
    \end{tikzcd}
\end{equation}

\end{document}