### 2023-10-13 Presentations/lightning_slides.tex
new pitch slides

\documentclass{beamer}
\usepackage[utf8]{inputenc}

\usepackage{amsmath,amsfonts,amssymb,amsthm,
mathtools,mathrsfs}
\usepackage{physics}
\usepackage{xcolor}
\usepackage{hyperref}

\usepackage{tikz-cd} % arrow diagram

% Citation style
\usepackage[style=apa, backend=biber, natbib]{biblatex}
\addbibresource{../references.bib}

\usetheme{Madrid}
\usecolortheme{default}
\useoutertheme[subsection=false]{miniframes}
\setbeamertemplate{frametitle continuation}{}
\setbeamertemplate{bibliography item}[triangle]
\setbeamertemplate{sidebar right}{}
%\addtobeamertemplate{footnote}{}{\vspace{2ex}}

% Presenter notes
%\setbeameroption{show notes}

\title[Optimal Cryptoeconomic Policies]{Dynamic Modeling for Optimal Cryptoeconomic Policies}
\subtitle{Pitch Proposal Fall 2023}
\author{Mingxuan He}

\institute[UChicago]{
M.A. in Computational Social Science -- Economics\\
Department of Economics, University of Chicago\\
mingxuanh@uchicago.edu
}

\date{\today}

\begin{document}

% title page
\begin{frame}
\titlepage  
\end{frame}

% table of contents
% \begin{frame}
% \frametitle{Table of Contents}
% \tableofcontents
% \end{frame}

%----------------
\section{Introduction}

\begin{frame}{Research Question}

\textbf{How should we design dynamic staking and burning policies for Proof-of-Stake cryptoeconomies?}\newline
\begin{itemize}
\item Current protocol-coded policies are static even though crypto-economies face dynamic shocks. ``Staking" and ``burning" rates are the primary coded policy rules in the underlying blockchain system
\begin{itemize}
    \item \textbf{Staking (increases ``token supply", decreases ``tokens in circulation")}: mint new tokens and give to those who stake their existing tokens as interest. Similar to interest on reserves.
    \item \textbf{Burning (decreases ``token supply")}: burn (remove from circulation) a portion of the transaction fee paid by users during blockchain transactions
    \end{itemize} \textbf{}

    \item Policy goal: maximize welfare for \underline{good actors}: users (increase in token price) and validators (transaction fees)
\end{itemize}

\note{
shocks from the real economy: Banks failing, stock market, interest rates\\
shocks from other cryptoeconomies: Luna-Terra, FTX/Alameda
}

\end{frame}

%----------------
%\section{Literature Review}

\begin{frame}{Literature Review}
    \begin{itemize}
        \item Micro foundations:\\
        {\footnotesize \textit{\citet{nisan2007algorithmic, budish2018economic, biais2019blockchain, gans2019more, gans2022mechanism, huberman2021monopoly}, etc.}}\\
        $\rightarrow$ \textbf{Block reward and transaction fee as incentive mechanisms}

        \item Pricing models of cryptocurrencies:\\
        {\footnotesize Bitcoin: \textit{\citet{athey2016bitcoin, garratt2018bitcoin, schilling2019some, schilling2019currency, catalini2020some, biais2023equilibrium, bolt2020value, hinzen2022bitcoin, chiu2017economics}, etc.}}\\
        {\footnotesize Proof-of-stake: \textit{\citet{catalini2020markets, saleh2019volatility, saleh2021blockchain}}}\\
        $\rightarrow$ \textbf{Static token supply means all shocks are absorbed by price}
        \item Monetary policies for stablecoins, platform tokens, and CBDCs:\\
        {\footnotesize \textit{\citet{cong2021tokenomics, cong2022token, cong2022staking, d2022can, fernandez2021central, zhu2019framework, sockin2023model, sockin2023decentralization}, etc.}}\\
        $\rightarrow$ \textbf{Optimizing policy for defending peg and/or making profit}
    \end{itemize}
\end{frame}

\begin{frame}{Contribution to the Literature}
\begin{itemize}
    \item Practical: Most cryptoeconomic systems feature fixed-schedule token supply
    \begin{itemize}
        \item deterministic token supply (Bitcoin\footnote{Satoshi Nakamoto (2008). Bitcoin: A peer-to-peer electronic cash system}), 
        fixed burn rate (Ethereum\footnote{Ethereum Documentation. ``Gas and Fees - Base fees"}), 
        naive dynamic burning (Binance Coin)
    \end{itemize}

    \item No published literature has been established on optimal staking and burning policies for proof-of-stake cryptocurrencies.
\end{itemize}
Novelty of this research:
\begin{itemize}
    \item Applies dynamic general equilibrium methods to model and optimize crypto policies for staking and burning            
      \item Traditional models for optimal fiscal policy (e.g. Ramsey) and monetary policy (e.g. NK) are not directly transferrable to cryptoeconomies due to transaction fee role in the cryptoeconomy

\end{itemize}
\end{frame}

%----------------
\section{Model}
% \begin{frame}{Inspiration}
%     \begin{figure}
%         \includegraphics[width=0.75\textwidth]{biais_title.png}
%     \end{figure}
%     * Bitcoin's economic/consensus system (proof-of-work) is different from proof-of-stake
% \end{frame}

\begin{frame}{A model of proof-of-stake cryptoeconomies}
    \[
\begin{tikzcd}[ampersand replacement=\&, column sep = 4.5em, row sep = 4.5em]
Protocol 
\ar[d, shift right, "\substack{\text{newly minted}\\\text{tokens}\\(t+1)}" ']
\&
\ar[d, shift right, "\substack{\text{endow-}\\\text{ment}}" 'near start]
\ar[d, shift left, "\substack{\text{purchase}\\ \text{assets}}" near start]
\&
\ar[d, "\substack{\text{transact.}\\ \text{benefits}}" ]
\&
\\
Validators
\ar[u, shift right, red, "\substack{\text{stake}\\\text{tokens}\\(t)}" ' red ]
\& 
\substack{Young\\Users}
\ar[l,"\substack{\text{variable}\\ \text{fees (``tip")}}" ']
\ar[lu, red, "\substack{\text{base fees}\\ \text{(burned)}}" {sloped,auto}' red]
\ar[r, dashed, "t+1", 
"\substack{\text{hold}\\\text{assets}}" ']
\& 
\substack{Old\\Users}
\ar[r, "\substack{\text{hacked}\\ \text{tokens}}"]
\&
Hackers
\end{tikzcd}
\]

\begin{itemize}
    \item Inspired by Biais et al. (J Finance 2023) ``Equilibrium Bitcoin Pricing"
    \item Two-period OLG model
    \item Endogenous token supply (innovation over \citet{biais2023equilibrium})
    \item Assumption: Crypto's fundamental value comes from the stream of future transactional benefits\\
    \begin{itemize}
        \item e.g. access to unique goods, not expropriated/taxed/constrained by government, direct internet access
    \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{User's problem}
    \begin{itemize}
    \item Period t: allocates endowment income $e_t^u$ into three asset classes: risk-free $s_t$, crypto $q_t$, and fiat $\hat{q}_t$. To buy crypto they pay a two-part transaction fee $\varphi_t^b$ (burned) and $\varphi_t^v$ (goes to validator)\\
    \item Period t+1: receive interest $r_t$, transactional benefits $\theta_{t+1}$, and hacking loss $h_{t+1}$\\
    \end{itemize}

    \begin{align}
    c_t^u &= e_t^u - s_t - (1+ \varphi_t^b+\varphi_t^v) q_t p_t - \hat{q}_t \hat{p}_t\\
    c_{t+1}^u &=  s_t (1+r_t) + (1-h_{t+1})(1+\theta_{t+1}) q_t p_{t+1} + \hat{q}_t \hat{p}_{t+1}
    \end{align}

\end{frame}

\begin{frame}{Validator's problem}
    \begin{itemize}
        \item Period t: receive transaction fees $\varphi_t^v$, stake $L_t$
        \item Period t+1: receive staking yield $\delta_t$
    \end{itemize}
    \textbf{Trade-off:}
    \[
    \text{more staking at } t \Rightarrow
    \begin{cases}
        \text{more staking yield at } t+1\\
        \text{less circulating supply} \Rightarrow \text{less transact. fees at } t
    \end{cases}
    \]
    \begin{align}
    c_t^v &= e_t^v + \varphi_t^v q_t p_t - L_t p_t \\
    c_{t+1}^v &= (1+\delta_t) L_t p_{t+1} 
    \end{align}
\end{frame}

\begin{frame}{Token supply}
    \begin{align}
        M_{t+1} &= (M_t - L_t) + L_t (1 + \delta_t) - \varphi_t^b X_t \nonumber \\
        &= M_t + \delta_t L_t - \varphi_t^b X_t
    \end{align}
    $M_t$: total token supply\\
    $X_t \equiv M_t - L_t$: circulating supply \\
    \quad $\Rightarrow$ Token market clearing condition is $X_t=q_t$\\
    \ \\
    $^*$ In reality the total reward is $\delta_t L_t^\alpha$, with pre-set $\alpha\in [0,1]$\footnote{$\alpha=0.5$ for Ethereum. See Ethereum Documentation. ``Proof-of-Stake Rewards and Penalties"}. Here I set $\alpha=1$ WLOG since $\delta_t$ is dynamic.
    % calculations: https://blog.bitmex.com/ethereums-proof-of-stake-system-calculating-penalties-rewards/
\end{frame}

\begin{frame}{First order conditions}
    * for a deterministic version of the model
    \bigskip

    User's Euler equations:
    \begin{equation}
        \frac{u'(c_t^u)}{u'(c_{t+1}^u)} = \beta (1+r_t) 
        = \beta \frac{(1+\theta_{t+1})(1-h_{t+1})}{1+\varphi_t^b+\varphi_t^v} \frac{p_{t+1}}{p_t}
        = \beta \frac{\hat{p}_{t+1}}{\hat{p}_t}
    \end{equation}

    Validator's Euler equation:
    \begin{equation}
        \frac{u'(c_t^v)}{u'(c_{t+1}^v)} = \beta (1+\delta_t) \frac{p_{t+1}}{p_t}
    \end{equation}

    WIP: Equilibrium solutions with stochasticity \& log/CRRA preferences
\end{frame}

%----------------
\section{Calibration}
\begin{frame}{Calibration (idea phase)}
    \begin{itemize}
        \item \textbf{Data sources:} Ethereum prices from Messari API, txn fees from Dune Analytics, staking data from beaconcha.in
        \item \textbf{Period:} Sep 2022 - now, down to block frequency (12s) 24/7
        \item Proxy for transactional benefits/convenience yields: average onchain txn processing per second $a_t$\citep{cong2022staking}
        \begin{itemize}
            \item other potential proxies: Metcalfe's measure $\log(DailyActiveAddresses^2)$; \href{https://www.ncsl.org/financial-services/cryptocurrency-2023-legislation}{index on US crypto regulation (NCSL.org)}
        \end{itemize}

        \item Proxy for hacking: manually index major hacks and wallet losses
    \end{itemize}
\end{frame}

% \begin{frame}{More extensions to model \& estimation}

% \begin{enumerate}
%     \item Modify validators' budget constraint: incorporates costs of validation under PoS and PoW \parencite{biais2019blockchain, saleh2019volatility, saleh2021blockchain}
%     \item Improvement in model calibration 
%     \begin{itemize}
%         \item novel time series data on Ethereum price and transaction fees
%         \item estimate transactional benefits using the NVT (network value to transaction ratio) or RVT (realized value to transaction ratio) metric instead of event-based index
%     \end{itemize}
%     \item Use vector autoregression models for exogenous process in variable transaction fees and transaction benefit (e.g. sign restrictions)
% \end{enumerate}
% \end{frame}

% \begin{frame}{Proposed Project Timeline}
% \begin{itemize}
%     \item Spring '23: As final project for ECMA33603 (Macro \& Financial Frictions), replicate the baseline model and implement extensions part 1 (endogenous token supply)
%     \item Summer '23: Implement extensions part 2; gather ideas from industry internship in Ethereum research \& dev
%     \item Fall '23 - Winter '24: Extensions part 3: Data collection, model calibration (likely using Ethereum data)
%     \item Spring '24: Finish paper
% \end{itemize}
% \end{frame}

% -----------------
\section*{}
\begin{frame}[allowframebreaks, noframenumbering]{References}
    %\nocite{*}
    %\bibliographystyle{apalike}
    %\bibliography{references}
    \printbibliography
\end{frame}

\end{document}